---
title: "STATS 782 Assignment 1; University of Auckland"
date: 'Due Date: 10:00 NZ Time, Monday 20 March 2023'
output:
  html_document: default
  pdf_document: default
---

### "Junyi Yang", "223831196"

I have read the declaration on the cover sheet and confirm my agreement with it.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) ## makes sure the code is shown
```


## Question 1

### a)

```{r}
seq(0,6,by=1.5)
```

### b)

```{r}
cumsum(seq(2,20,by=2))
```

### c)

```{r}
abs(3:-3)
```

### d)

```{r}
rep(seq(40,0,by=-10),times = 4) + rep(0:4,each=4)
```

### e)

```{r}
paste(LETTERS[7:1],7:1,'-',letters[1:7],1:7,sep = '')
```



## Question 2

```{r}
## define f here
f = function (x) {
  
  # we also need to return the empty and 
  # length 1's situation 
  if (length(x)<=1) {
    return(numeric(0))
  }
  
  
  # this is the normal part
  # get the a's next minu the before one
  a = which(diff(x) >= 0)
  if (length(a)==0){
    return(x)
  }
  
  # special case
  if (length(a)==1){
    # we need the head and last
    a = c(0,a,length(x))
    
  }else{
    if (x[length(x)] < x[length(x) - 1]) {
      a = c(a, length(x))
    }
  }
  
  # if the diff is 1 , then there is no elements between them
  check = (diff(a) [which.max(diff(a))] == 1)
  if (check){
    return(numeric(0))
  }
  start_index = which.max(diff(a))
  end_index = start_index + 1
  # plus 1 for the we put next to min the before one
  x[(a[start_index] + 1):a[end_index]]
}


```

```{r}
## uncomment the tests when you are done
x <- c(-1, 2, 7, 5.5, 5, 4, 4, 3, 2, -4, -4.1, -5, 8)

f(x)

set.seed(782)
f(0)
f(3:1)
f(1:3)
f(c(Inf, 0, -Inf, pi, 3))
f(round(runif(100) * 100, 1))
```


## Question 3

### a)

```{r}
# at first we need to do the preparation for the data 
library(MASS)

state <- rownames(road) # name of the US state
deaths <- road$deaths # number of deaths
drivers <- road$drivers # number of drivers (in 10,000s)
temp <- road$temp # average daily max temperature in January
```

#### We need a new chunk to create the main part:

```{r}
max(deaths)
```


### b)

```{r}
mean(deaths)
```

### c)

```{r}
# to get this result we need to get
# the fatality rate = number of deaths / number of drivers

fatality_rate = deaths / drivers

# then get the index for the max
index = which.max(fatality_rate)

# then we get the name 
state[index]
```

### d)

```{r}
# at first we need to get he average number of deaths
avg = mean(deaths)

# then we get the close rank for the deaths R
close = state[order(abs(deaths-avg))][1:3]
close
```

### e)

```{r}
# if we want the at least 1,000,000 drivers
# we need to change get the drivers' items

over_drivers_index = which((drivers / 100)>=1)

# then we calculate the highest fatality rates
fatality_rate_per = deaths[over_drivers_index] / (drivers*100)[over_drivers_index]

# then we get 3 the highest rates
state[over_drivers_index][order(fatality_rate_per,decreasing = TRUE)][1:3]


```


## Question 4

### a)

```{r}
# for we only need the first 7 terms of the series

# in order to get the more accurate result we do not use vector numbers for the denominator

# our arctan function will be 
arctan_7 = function(x) {
  arct = x - (x^3)/3 + (x^5)/5 - (x^7)/7 + (x^9)/9 - (x^11)/11 + (x^13)/13
  arct
}


# then we get the value for pi
4*arctan_7(1)
# and the pi in R
pi

# difference
diff = abs(4*arctan_7(1) - pi)
diff
```

Clearly this is not a good approximation. The difference between our result and 'pi' which is constant in R is 0.1421458 which is big.


### b)

```{r}
## define f here

f = function(n) {
  if(n == 1){
    return (1)
  }
  
  num = (2*seq(1,n)-1)
  # we use 4*arctan(1) = pi
  # so the x value is 1
  arctan = as.double(sum((c(1,-1) * 1^num) / (num))) 
  # return value
  4*arctan
}
# f(1000)
f(1000)
```

### c)

```{r}
## define approx here
approx = function(eps){
  # we create a Enough number to be the start precision
  result = as.double('Inf')
  # we start from n is 1
  current = 1
  
  # use app to store all the result
  app = result
  
  # we also need a flag 
  # if we have it to be TRUE 
  # means we have in the dead loop
  flag = 0
  
  # we want to compare the result_i with the result_(i-1) 
  # if their difference is lower than the 2*eps 
  # we consider that we get the right approximate for the  value 'pi'
  while (result > 2*eps) {
    if (current == 1){
      current = current + 1
      next
    }
    
    
    app_i = f(current)
    app_i_before =  f(current-1)
    result = abs (app_i - app_i_before)
    
    current = current + 1
    
  }
  c(f(current-1),current-1)
  
}
```

### Why:

1. What we want is the approximation value for 'pi' which has the approximation error is less than **eps**, and the number of elements of the series you used. 

2. For we have the approximation function 'f(n)' in the question **(b)**, we can use this function to do get n from 1 to some value m.

3. We get this 'm' value when the value abs(f(m) - f(m-1)) < 2*eps as the convergence condition. 
Because we take 2 approximations f(m) and f(m-1), if we consider them to be the convergence condition, each of their difference with the real value 'pi' should be less than eps.

4. But we do not know the value for the real 'pi'. So we use their difference. If they all of their difference with the real value 'pi' are less than eps, the difference between f(m) and f(m-1) must be less than 2*eps.

5. So after about 100,000 times increase for n from 1. We get that the f(100001) is the approximation which we want. And the number of elements of the series we used is 100,000 


```{r message=FALSE, warning=FALSE}
# approx(0.00001)

# use this for we have no same length in the f function
approx(0.00001)
```




### d)

```{r}
## define sr here

sr = function(n){
  i = 0:(n-1)
  part2 = 
    (factorial(4*i) * (1103 + 26390*i)) / ((factorial(i)^4) * 396^(4*i))
  
  
  part2
  
}

# sr(n = 4)
sr(n=4)
# compare and explain
part1 = (2*sqrt(2) / 9801 )
# then our pi will be
result =  part1*sum(sr(n=4))
our_pi = 1 / result

abs(our_pi-pi)
```

From the result, we could know that we get a very good approximation of pi value. Our accuracy has exceeded the range we can express which is that we have the '0' in our difference result without any decimal digits. 

### e)

```{r}
# we just need to compose the n = 20
our_pi_new = 1/(part1 * sum(sr(20)))

abs(our_pi_new-pi)
```

### Why :

1. From the result, we could see that our absolute value of difference can not be small (We cannot express it as a smaller number). The result get from sr(n=20) is same from the result sr(n=4). 

2. For the precision of R to calculate is about 16 decimal digits, and these 2 results from the function are less than 16 decimal digits.

3. So we have the same result for these 2 (all difference with pi is 0).

### How many decimal:

1. As what we have mentioned before the precision of R to calculate is about 16 decimal digits, we can choose 16 decimal which we expect to be correct.



### f)

```{r}
sr(50)

result = 1/ ( part1 *sum(sr(50)))
result
```

### Explain in detail why:

1. This is because we have the 'NaN' value in sr(50) function. So the final approximation will be 'NaN'.

### Root cause:

1. Root cause is that the number we can represent has a limited range in R. Once the value of a number is greater than the range we can represent, the number will be expressed as 'Inf'. If we divide 'Inf' by 'Inf', we will get a value 'NaN'.

2. In our 'sr(n)' function, the growth rate of the denominator is greater than the numerator, so the denominator reaches the 'Inf' value first. So we start with n=23, and the value of sr (n) is 0. After that, our molecule also increased to the value of 'Inf' (n is about 43). We divide the numerator by the denominator to get the 'NaN' value
