---
title: "lab03"
author: "Junyi Yang"
date: "2023-03-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1

```{r}
m = matrix(1:9, 3)
```

```{r}
m[2, 3]
```

```{r}
# this code will delete the second line and the first col
m[-2,-1]
```

```{r}
m[2:6]
```

```{r}
# will be m[3,3],m[3,2]
# and m[2,3] and m[2,2]
# the nmake it to a matrix
m[3:2,3:2]
```

```{r}
# bind the matix with 10 
# use the row bind
rbind(10, m)
```

```{r}
# calculate the sum for each col
colSums(m)
```

```{r}
# rep 2:4 to add to each col
# like 2,3,4,2,3,4,2,3,4
m + 2:4
```

```{r}
# repeat for each 
# like 2,2,2,3,3,3,4,4,4
m + rep(2:4, each=3)

```

```{r}
 m[row(m) < col(m)]
```

## Q2

```{r}
set.seed(782)

(x = runif(5))     # a uniform random sample

mean(x)            # mean

sd(x)              # standard deviation

(z = (x - mean(x)) / sd(x)) # standardised sample

mean(z)            # why not exactly zero?

round(mean(z), 10) # now exactly zero

sd(z)
```

#### Why not Zero?

1. For we use some close value to get the x - mean(x) value, we will have some lack of precision. So the result mean is not zero.

2. And we have the small sample size, this may cause the result to have some difference as we expected.



```{r}
set.seed(782)
m = matrix(runif(15), nrow=5)
colnames(m) = paste0("X", 1:ncol(m))

m
```

### 1) Write R code using for loop (not apply and friends yet) which standardises each column of m and creates a resulting matrix z. You can start with z = m and then modify the columns.

```{r}
z = m
# we need to do the standardises for each cols
for (j in 1:ncol(z)) {
  mean = mean(z[,j])
  sd = sd(z[,j])
  sd_z = (z[,j] - mean) / sd
  z[,j] = sd_z
  print(round(mean(z[,j]), 10))
  print(sd(z[,j]))
}
z
```

The result seams great!

### 2)Use apply() to find all the column means and standard deviations of m, and save them in variables mu and sig, respectively. Apply the same to the standardised z computed above to verify that it is correct without storing the result.

```{r}
# we need to init the z
z = m

# colmeas
mu = apply(z,2,mean)
mu

# stadard deviations
sig = apply(z,2,sd)
sig

# check 
for (j in 1:ncol(z)) {
  mean = mean(z[,j])
  print(mean)
  sd = sd(z[,j])
  print(sd)
}
```

```{r}
# standard 


```

